Work Fragmentation in Developer Interaction Data

### 2016-11-08 paper_2.tex
Removed the blue highlight
typos and table XV description

\setboolean{showchangedforreviewer}{true} % toggle to show or hide sentences and paragraphs requested by the reviewers.

@@

Second, we have kept all traces without aggregate information. Unfortunately, we cannot know the distribution in time of the actions that have been aggregated in a single event: without additional information, an aggregation of 15 events over one hour is as likely every 4 minutes as it is to have 10 events in the first 5 minutes and 5 more in the last 10 minutes. Therefore, designing a correct disaggregation task to handle these traces traces is very difficult, if not impossible.

@@

\begin{changedforreviewerlong}
To ensure that a small subset of users do not dominate our dataset, we counted the number of sessions attributed to each UDC user. The third quartile was 5 sessions, the median 15, and the third quartile 38; the minimum was 1, and the maximum 270. These figures tell us that if some users have more sessions than others (which is to be expected), there is no major bias due to a small number of users concentrating the vast majority of the sessions.
\end{changedforreviewerlong}

@@

Following the conclusions on RQ1, the number of interruptions has an impact on the metrics and should not be ignored. Further, as the number of interruptions increases the difference between the groups minimizes.  %we believe that having   more distinctive groups instead of groups created by ranges is also valid.
 Thus we want to compare the proportions of prolonged interruptions among sessions that have the \emph{same} number of interruptions. We hence split the \emph{low} and \emph{high} groups in subgroups of sessions which have the same number of interruptions. We consider groups of up to seven sessions, as larger groups are small. We finally compare pairs of \emph{low} and \emph{high} groups that have the same number of interruptions, expecting higher productivity indicators in the \emph{low} groups.

@@

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/all_ints} \\
\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/udc_all_ints_2} 
\end{tabular}
\caption{Local effect of an interruption in the user activity for Mylyn (top) and UDC (bottom)}
\label{fig:all_ints}
\end{figure}

@@

\begin{table}[ht!]
	\small
	\renewcommand{\arraystretch}{1.3}
	\caption{New category of events. }
	\label{tbl:new_events_udc}
	\centering
	\begin{tabular}{|p{2.5cm}|p{9 cm}| p{2cm} |}
		\hline 
		category & description & type \\  
		\hline 
		\hline 
		high-nav & High-level navigation through classes, windows and artifacts & navigation \\
		file & Management of files (saving, closing, opening, etc.) & navigation\\
		edit-text & Text edition events & edition \\
		debug & Events executed during debugging sessions & navigation \\
		text-nav & Low-level text navigation & edition \\
		search & Text or code search & navigation \\
		find-replace & Search and replacement of text or code entities & navigation\\
		perspective & Change of perspectives in Eclipse & navigation  \\
		control & Source control events & navigation \\
		refactoring & Refactoring events & edition \\
		clean-build &  Events related to the build system (clean, build or run a project)  & navigation \\
		\hline
	\end{tabular}
\end{table}

@@

To get a baseline, the same frequency calculation was made for the entire dataset, without distinction of the location of events in the session. We then compared the weighted average changes to have evidence of what kind of events are more often executed after a positive or a negative interruption. The Table \ref{tbl:stats_events} shows the results, and is explained below.

@@

The results presented in this work provides of evidence of the effect of work fragmentation observed in interaction data that is in accordance to what observational studies had concluded before. We began by comparing quantities (number and length of interruptions) and working patterns. The related work to this approach suggests that as a consequence to interruptions there is a recovery time that demands extra time to resume the task, so we expected to see a decrease in the metrics as the number of interruptions increases \cite{MGH05,IH07,CHW04}.

@@

\paragraph{Comparison with observational studies}
Our theoretical foundations are based on observational studies performed on information workers' development environments \cite{MGH05, IH07, LVD06, PD10}. These findings helped us understand the response to work fragmentation according to the recorded user activity, and it was possible to made a relation between the user's data and the phenomena that occur on the working places described in the studies.

@@

The workbench activation and deactivation events were a particular issue, for they can be executed under many circumstances. We performed live test to understand how these events work and gain confidence about their precise behavior. For instance, we found that they were reliable markers of when the user leaves and returns to Eclipse---and we used this fact to accurately detect interruptions---but we had more uncertainty on other scenarios. Therefore, we decided not count them as ``system'' events, and thus only use them to define an interruption or the end of a session.

>>>>

\setboolean{showchangedforreviewer}{false} % toggle to show or hide sentences and paragraphs requested by the reviewers.

@@

Second, we have kept all traces without aggregate information. Unfortunately, we cannot know the distribution in time of the actions that have been aggregated in a single event: without additional information, an aggregation of 15 events over one hour is as likely every 4 minutes as it is to have 10 events in the first 5 minutes and 5 more in the last 10 minutes. Therefore, designing a correct disaggregation task to handle these traces is very difficult, if not impossible.

@@

\begin{changedforreviewerlong}
To ensure that a small subset of users do not dominate our dataset, we counted the number of sessions attributed to each UDC user. The first quartile was 5 sessions, the median 15, and the third quartile 38; the minimum was 1, and the maximum 270. These figures tell us that if some users have more sessions than others (which is to be expected), there is no major bias due to a small number of users concentrating the vast majority of the sessions.
\end{changedforreviewerlong}

@@

Following the conclusions on RQ1, the number of interruptions has an impact on the metrics and should not be ignored. Further, as the number of interruptions increases the difference between the groups minimizes.  %we believe that having   more distinctive groups instead of groups created by ranges is also valid.
 Thus we want to compare the proportions of prolonged interruptions among sessions that have the \emph{same} number of interruptions. We hence split the \emph{low} and \emph{high} groups in subgroups of sessions which have the same number of interruptions. We consider groups of up to seven sessions, as larger groups are in small amount. We finally compare pairs of \emph{low} and \emph{high} groups that have the same number of interruptions, expecting higher productivity indicators in the \emph{low} groups.

@@

\begin{figure}[!ht]
\centering
\begin{tabular}{c}
\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/all_ints} \\
\includegraphics[width=0.6\linewidth,clip=, angle=0]{figures/udc_all_ints_2} 
\end{tabular}
\caption{Local effect of an interruption in the user activity.}
\label{fig:all_ints}
\end{figure}

@@

\begin{table}[ht!]
	\small
	\renewcommand{\arraystretch}{1.3}
	\caption{New classification of events that reflects in detail the nature of each event. }
	\label{tbl:new_events_udc}
	\centering
	\begin{tabular}{|p{2.5cm}|p{9 cm}| p{2cm} |}
		\hline 
		category & description & type \\  
		\hline 
		\hline 
		high-nav & High-level navigation through classes, windows and artifacts & navigation \\
		file & Management of files (saving, closing, opening, etc.) & navigation\\
		edit-text & Text edition events & edition \\
		debug & Events executed during debugging sessions & navigation \\
		text-nav & Low-level text navigation & edition \\
		search & Text or code search & navigation \\
		find-replace & Search and replacement of text or code entities & navigation\\
		perspective & Change of perspectives in Eclipse & navigation  \\
		control & Source control events & navigation \\
		refactoring & Refactoring events & edition \\
		clean-build &  Events related to the build system (clean, build or run a project)  & navigation \\
		\hline
	\end{tabular}
\end{table}

@@

To get a baseline, the same frequency calculation was made for the entire dataset, without distinction of the location of events in the session. We then compared the weighted average changes to have evidence of what kind of events are more often executed after a positive or a negative interruption. The Table \ref{tbl:stats_events} shows the results, and it is explained below.

@@

The results presented in this work provides some evidence of the effect of work fragmentation observed in interaction data that is in accordance to what observational studies had concluded before. We began by comparing quantities (number and length of interruptions) and working patterns. The related work to this approach suggests that as a consequence to interruptions there is a recovery time that demands extra time to resume the task, so we expected to see a decrease in the metrics as the number of interruptions increases \cite{MGH05,IH07,CHW04}.

@@

\paragraph{Comparison with observational studies}
Our theoretical foundations are based on observational studies performed on information workers' development environments \cite{MGH05, IH07, LVD06, PD10}. These findings helped us understand the response to work fragmentation according to the recorded user activity, and it was possible to make a relation between the user's data and the phenomena that occur on the working places described in the studies.

@@

The workbench activation and deactivation events were a particular issue, for they can be executed under many circumstances. We performed live test to understand how these events work and gain confidence about their precise behavior. For instance, we found that they were reliable markers of when the user leaves and returns to Eclipse---and we used this fact to accurately detect interruptions---but we had more uncertainty on other scenarios. Therefore, we decided not to count them as ``system'' events, and thus only use them to define an interruption or the end of a session.
